Financial markets for reaching ``green" targets: Cap-and-trade strategies in the Chilean electricity sector

### 2020-04-11 draft_paper/draft_v1.tex
Updates from Overleaf

\section{Introduction}\label{sec:intro}
Over 150 years ago, Eunice Foote, in her paper ``\textit{Circumstances affecting the Heat of the Sun's Rays}'' \cite{Foote1856} discovered for the first time in history the link between temperature and CO$_2$ and recognized the effect this link might have on the atmosphere: ``\textit{An atmosphere of that gas would give to our earth a high temperature; and if as some suppose, at one period of its history the air had mixed with it a larger proportion than as present, an increased temperature from its own action as well as from increased weight must have necessarily resulted}''. This work and other groundbreaking studies that came afterwards set the basis of climxate science. To date, the unanimous scientific consensus states that Earth is warming and that the main cause are the anthropogenic Greenhouse gases (GHG) emission \cite{powell2019scientists}. One of the main ``culprits'' of polluting the environment is the energy sector, burning fossil fuels to produce heat and electricity. In fact, driven by higher energy demand in 2018, global energy-related CO$_2$ emissions reached a historic high of 33.1 GtCO$_2$ \cite{IEA2019}. Hence, the energy sector appears as a key participant in the fight towards climate crisis mitigation. This is one of the reasons why energy-sector's decarbonization strategies have been the focus of a vast number of studies, being carbon pricing the main mechanism to accomplish such a goal \cite{Robert2019Carbon}.

@@

Mu\~{n}oz, Pumarino \& Salas \cite{munoz2017aiming} commented in their conclusions that in some cases, regulators do not performed careful analysis of the impact of policies that are implemented in their territories. In this sense, our paper is a contribution precisely in the topic of a cap-and-trade system for Chile, a system that Chile's government has manifest its interest as the Roadmap document shows, but has not performed deep analysis of its implementation or impact in their environmental policies. We model the Chilean power market including a cap-and-trade system using a two-staged capacity expansion equilibrium model \cite{ehrenmann2011stochastic}. The analysis considers a stochastic equilibrium, since we model the present operation and investment for an uncertain future, in a perfect competition market. We studied the effect of different carbon budgets (set as a cap) in the electricity production mix in Chile. The cap is estimated as the remaining carbon budget that is allowable in the electric market. We show that the emission target in Chile's NDC is not restrictive by itself, and does not change substantially the technology mix. Also, we found that the price of the carbon allowances, compared with the current tax, is much more higher and more in agreement with what previous studies have shown. In the next sections we describe the model and the assumptions made (Section \ref{sec:model}); then we give details of the Chilean electricity market and the data used (Section \ref{sec:chiledata}, and we present the results of our analysis and final remarks in Sections \ref{sec:scenarios} and \ref{sec:concl}.

@@

Constraint~\ref{eq:c1} represents the generation capacity constraint for each producer $i$ in scenario $\omega$ and time $t$ in the second stage. Similarly, Constraint~\ref{eq:c2} restricts generation for $t=0$ (first stage). Additionally, we limit the total capacity (installed and new investments) for producer $i$ to the resource potential for each technology, as shown in Constraint~\ref{eq:c3}. The maximum amount of allowances that each generator can sell is limited by CO$_2e$ permits bought from the auctioneer (Constraint~\ref{eq:c4}) and the total emissions of each producer is restricted to all the transactions of allowances in the market (Equation~\ref{eq:c5}). Variables in parenthesis next to each constraint represent the respective Lagrange multipliers.

@@

For each generation technology, we obtained the data for capital investment and operation cost (CAPEX and OPEX), installed capacity and future planned capacities expansion through the \textit{Comisi\'on Nacional de Energ\'ia} (CNE) website. Information about the resource potentials were obtained from \cite{Santana2014potencial}.
In the case of the demand, we also use projection provided the CNE\footnote{http://datos.energiaabierta.cl/dataviews/251663/proyeccion-de-demanda-en-gwh/}, which foresees the demand until year 2038. This data is linearly extrapolated to extend the demand projection up to 2050. \pia{This demand  is the one used in the deterministic analysis.}

@@

The model is calibrated to the above data, considering the operation in 2018. This calibration was performed by fine-tuning a subset of the parameters of the model, such as cost and the technology/investment change, $TC$ and $TCR$ (see Section~\ref{sec:model} for the details). In any case, the estimation of these parameters follows the trends expected for future years (as presented by \cite{Mena2019csp,NRELcosts}), where investment and operation costs of NCRE sources are expected to decrease, as opposed to fuel-based generation where costs tends to remain constant \cite{Mena2019csp,NRELcosts}.
Figure~\ref{fig:bc}(a) shows the electric generation mix produced in our model for 2018. Most of the electricity production of the country comes from coal and gas-fueled generators. Only 18\% of the energy matrix comes from NCRE sources, and taken into account large hydropower plants, this percentage increases up to  46\%. This values are in good agreement with the real operation data from \textit{Comisi\'on Nacional de Energ\'ia} \cite{CNE2018}, which estimates that the contribution from NCREs is 20\% and raises to 50\% when large hydropower plants are included.

@@

\begin{figure}[ht]
    \begin{subfigure}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{draft_paper/Figures/demand_scenarios.png}
    \caption{}
  \label{fig:demand_scn}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{draft_paper/Figures/mean_prices_scenarios.png}
    \caption{ }
  \label{fig:mean_price}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{draft_paper/Figures/price_allowances.png}
    \caption{  }
  \label{fig:price_allo}
  \end{subfigure}
    \begin{subfigure}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{draft_paper/Figures/prices_trading_scenarios.png}
    \caption{}
  \label{fig:price_trade}
  \end{subfigure}\quad
   \caption{Panel (a): Demand scenarios. Panel (b): Electricity prices per MWh, for each carbon budget value and different demand scenarios. Color legend same as in Panel (a). Panel (c): Price of allowances per ton CO$_2e$ for each carbon budget. Panel (d): Allowance trading price per ton CO$_2e$ for each carbon budget and different demand scenarios. Color legend same as in Panel (a).}
\end{figure}

>>>>

\section{Introduction}\label{sec:intro}
Over 150 years ago, Eunice Foote, in her paper ``\textit{Circumstances affecting the Heat of the Sun's Rays}'' \cite{Foote1856} discovered for the first time in history the link between temperature and CO$_2$ and recognized the effect this link might have on the atmosphere: ``\textit{An atmosphere of that gas would give to our earth a high temperature; and if as some suppose, at one period of its history the air had mixed with it a larger proportion than as present, an increased temperature from its own action as well as from increased weight must have necessarily resulted}''. This work and other groundbreaking studies that came afterwards set the basis of climate science. To date, the unanimous scientific consensus states that Earth is warming and that the main cause are the anthropogenic Greenhouse gases (GHG) emission \cite{powell2019scientists}. One of the main ``culprits'' of polluting the environment is the energy sector, burning fossil fuels to produce heat and electricity. In fact, driven by higher energy demand in 2018, global energy-related CO$_2$ emissions reached a historic high of 33.1 GtCO$_2$ \cite{IEA2019}. Hence, the energy sector appears as a key participant in the fight towards climate crisis mitigation. This is one of the reasons why energy-sector's decarbonization strategies have been the focus of a vast number of studies, being carbon pricing the main mechanism to accomplish such a goal \cite{Robert2019Carbon}.

@@

Mu\~{n}oz, Pumarino \& Salas \cite{munoz2017aiming} commented in their conclusions that in some cases, regulators do not performed careful analysis of the impact of policies that are implemented in their territories. In this sense, our paper is a contribution precisely in the topic of a cap-and-trade system for Chile, a system that Chile's government has manifest its interest as the Roadmap document shows, but has not performed deep analysis of its implementation or impact in their environmental policies. We model the Chilean power market including a cap-and-trade system using a two-staged capacity expansion equilibrium model \cite{ehrenmann2011stochastic}. The analysis considers a stochastic equilibrium, since we model the present operation and investment for an uncertain future, in a perfect competition market. We studied the effect of different carbon budgets (set as a cap) in the electricity production mix in Chile. The cap is estimated as the remaining carbon budget that is allowable in the electric market. We show that the emission target in Chile's NDC is not restrictive by itself, and does not change substantially the technology mix. Also, we found that the price of the carbon allowances, compared with the current tax, is greater and consistent with what previous studies have shown. In the next sections we describe the model and the assumptions made (Section \ref{sec:model}); then we give details of the Chilean electricity market and the data used (Section \ref{sec:chiledata}), and we present the results of our analysis and final remarks in Sections \ref{sec:scenarios} and \ref{sec:concl}.

@@

Constraint~\ref{eq:c1} represents the generation capacity constraint for each producer $i$ in scenario $\omega$ and time $t$ in the second stage. Similarly, Constraint~\ref{eq:c2} restricts generation for $t=0$ (first stage). Additionally, we limit the total capacity (installed and new investments) for producer $i$ to the resource potential for each technology, as shown in Constraint~\ref{eq:c3}. The maximum amount of allowances that each generator can sell is limited by CO$_2e$ permits bought from the auctioneer (Constraint~\ref{eq:c4}) and the total emissions of each producer is restricted to all the transactions of allowances in the market (Constraint~\ref{eq:c5}). Variables in parenthesis next to each constraint represent the respective Lagrange multipliers.

@@

For each generation technology, we obtained the data for capital investment and operation cost (CAPEX and OPEX), installed capacity and future planned capacities expansion through the \textit{Comisi\'on Nacional de Energ\'ia} (CNE) website. Information about the resource potentials were obtained from \cite{Santana2014potencial}.
In the case of the demand, we also use projection provided the CNE\footnote{http://datos.energiaabierta.cl/dataviews/251663/proyeccion-de-demanda-en-gwh/}, which foresees the demand until year 2038. This data is linearly extrapolated to extend the demand projection up to 2050. This demand  is the one used in the deterministic analysis.

@@

The model is calibrated to the above data, considering the operation in 2018. This calibration was performed by fine-tuning a subset of the parameters of the model, such as cost and the technology/investment change, $TC$ and $TCR$ (see Section~\ref{sec:model} for the details). In any case, the estimation of these parameters follows the trends expected for future years (as presented by \cite{Mena2019csp,NRELcosts}), where investment and operation costs of NCRE sources are expected to decrease, as opposed to fuel-based generation where costs tends to remain constant \cite{Mena2019csp,NRELcosts}.
Figure~\ref{fig:bc}(a) shows the electric generation mix produced in our model for 2018. Most of the electricity production of the country comes from coal and gas-fueled generators. Only 18\% of the energy matrix comes from NCRE sources, and taken into account large hydropower plants, this percentage increases up to  46\%. This values are in good agreement with the real operation data from CNE \cite{CNE2018}, which estimates that the contribution from NCREs is 20\% and raises to 50\% when large hydropower plants are included.

@@

\begin{figure}[ht]
    \begin{subfigure}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{draft_paper/Figures/demand_scenarios.png}
    \caption{}
  \label{fig:demand_scn}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{draft_paper/Figures/mean_prices_scenarios.png}
    \caption{ }
  \label{fig:mean_price}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{draft_paper/Figures/price_allowances.png}
    \caption{  }
  \label{fig:price_allo}
  \end{subfigure}
    \begin{subfigure}[t]{0.5\linewidth}
    \includegraphics[width=\textwidth]{draft_paper/Figures/prices_trading_scenarios.png}
    \caption{}
  \label{fig:price_trade}
  \end{subfigure}\quad
   \caption{Panel (a): Demand scenarios. Panel (b): Electricity prices per MWh, for each carbon budget value and different demand scenarios. Panel (c): Price of allowances per ton CO$_2e$ for each carbon budget. Panel (d): Allowance trading price per ton CO$_2e$ for each carbon budget and different demand scenarios. Color legend for Panels (b) and (d) are the same as in Panel (a).}
\end{figure}
